---
title: "PML Course Project"
author: "Christian Müller"
date: "Saturday, January 24, 2015"
output: html_document
---
# Practical Machine Learning: Final Project
```{r, warning=FALSE}
library(caret) #Loading caret
set.seed(1234) #Setting seed for reproductability
```

## Data preparation 

The next lines will prepare the data: Firstly, it will clear up missing values. Secondly, variables with variance near zero or with A LOT OF missing values will be removed. Thirdly, we will delete time information. 

```{r, warning=FALSE}

tr <- read.csv("pml-training.csv")
prepare <- function(t) {
  t[t == ""] <- NA
  t[t == "#DIV/0!"] <- NA
  t <- t[-nearZeroVar(t)]
  t <- (t[,colSums(is.na(t)) < 1000 ])
  t <- t[-(which(names(t) %in% c("X","user_name","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp")))]
  t
}

tr <- prepare(tr)
head(tr)
```

## Splitting the data
Next step is splitting the data into my own training and testing set. Because of practicability and speed, I will use only 10 percent of the training set for machine learning. 

```{r, warning=FALSE}
inTrain_full <- createDataPartition(y=tr$classe, p=0.8, list=FALSE)
training_full <- tr[inTrain_full,]
testing <- tr[-inTrain_full,]
inTrain <- createDataPartition(y=training_full$classe,p=0.3, list=FALSE)
training <-  training_full[inTrain,]

```

## Random forest

Now, the data is ready for machine learning. I decided to use the random forest method, because random forest is in general a high performing algorithm and because some other data scientists used it in the past for Human Activity Recognition achieving good results (i.e Gjoreski et al.(2011)).

```{r, cache=FALSE, warning=FALSE}
modelFit <- train(training$classe~.,data=training ,method="rf",prox=TRUE)

print(modelFit)
print(modelFit$finalModel)
```

We achieved an accuray of more than 0.90% and an estimate of error rate of less than 0.05%. Thats a quite good result! Maybe we could achieve even more accuracy by using the whole training data set. But I will stay frugal. 

## Testing the model

Now, we will test the model with the test data.

```{r, warning=FALSE}
testingPC <- predict(modelFit, testing[-(which(names(tr) == "classe"))])
confusionMatrix(testing$classe , testingPC)
```

## Preparing submissions

In the following I prepare the submissions and the text file output.

```{r, warning=FALSE}
test <- read.csv("pml-testing.csv")
test <-  test[which(names(test) %in% names(training))]
answers <- predict(modelFit,newdata=test)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)

```
